---
title: "Reproducible Research: Peer Assessment 1 "
author: "Launia White"
date: "8/7/2020"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r lib, include=FALSE, echo=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
```

## Loading and preprocessing the data
```{r loaddata, echo=TRUE}
dat <- read_csv("~/activity.csv")
```
## What is mean total number of steps taken per day?
```{r mnsteps, echo=TRUE}
ts <- aggregate(steps~date, dat, FUN=sum)
hist(ts$steps, main="Histogram of Total Steps per Day", xlab="Total Steps")
options(scipen=999)
mnsteps <- round(mean(ts$steps), digits=1)
medsteps <- round(median(ts$steps), digits=1)
```

The mean total number of steps per day is `r mnsteps` and the median total number of steps taken per day is `r medsteps`.

## What is the average daily activity pattern?
```{r pattern, echo=TRUE}
ms <- aggregate(steps~interval, dat, FUN=mean, na.rm=TRUE)
p1 <- ggplot() + geom_line(aes(y = steps, x = interval),
      data = ms, stat="identity") + 
  ggtitle("Mean Steps per Day by Interval") + 
  labs(x="Interval", y="Mean Steps per Day") 
p1
maxsteps <- ms[order(-ms$steps),]
maxsteps <- head(maxsteps, n=1)
maxsteps$steps <- round(maxsteps$steps, digits=1)
```

Interval `r maxsteps$interval` had the maximum number of steps of `r maxsteps$steps`.

## Imputing missing values
```{r missing, echo=TRUE}
na <- sum(is.na(dat))
```

The total number of missing values in the dataset was `r na`.



```{r impute, echo=TRUE}
imputed <- dat # new dataset called imp
for (i in ms$interval) {
  imputed[imputed$interval == i & is.na(imputed$steps), ]$steps <- 
    ms$steps[ms$interval == i]
}
is <- aggregate(steps~date, imputed, FUN=sum)
hist(is$steps, main="Histogram of Total Steps per Day", xlab="Total Steps")
mnstepsi <- round(mean(is$steps), digits=2)
medstepsi <- round(median(is$steps), digits=2)
```
After imputation, the mean total number of steps per day was `r mnstepsi` and the median total number of steps taken per day increased to `r medstepsi`.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday, echo=TRUE}
imputed$daytype <- ifelse(weekdays(imputed$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
msdaytype <- aggregate(steps~interval + daytype, imputed, FUN=sum, na.rm=TRUE)
p3 <- ggplot() + geom_line(aes(y = steps, x = interval),
                           size=1, data = msdaytype, stat="identity") + 
  ggtitle(" ") + 
  labs(x="Interval", y="Number of steps") 
p4 <- p3 + facet_grid(daytype ~ .)
p4
```


